---
title: "FMWT drought stuff"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(emmeans)
library(visreg)
library(lme4)
library(lmerTest)
library(readxl)
```

## Longfin Smelt

I know Jereme doesn't like the annual FMWT index, but it's really easy to work with. I am also curious to see whether it tells a similar story to the Baesyian analysis of all the station data that Jereme is working up, so here goes.

I'll just grab the FMWT indexes and attach the water year types based on my drought/wet assignments and the Sac Valley index. 

```{r}
FMWT = read.csv("data/FMWTindices.csv")

#Year types are based on water year, whereas the FMWT is probably most senstative
#to the previous water year (so FMWT for Calendar Year 2020 is most sensative to water year 2020)
yeartypes = read_excel("data/Integrated data set.xlsx", sheet = "yearassignments")

FMWT = left_join(FMWT, yeartypes) %>%
  filter(Year > 1969)

#make a quick box plot
ggplot(FMWT, aes(x = Drought, y = Longfin.Smelt)) + geom_boxplot() + scale_y_log10()
```

## Data exploration

It probably isn't normally distributed, but probably not zero-inflated either.

```{r}
hist(FMWT$Longfin.Smelt)
hist(log(FMWT$Longfin.Smelt))
```

wow. That actually looks pretty nice.


# Now some models. 

First just log-transformed linear models. I know they are dorky, but they are also easy.

I'll do an ANOVA on drought/nutral/wet as well as a linear model on Sac valley index.

```{r}
l1 = lm(log(Longfin.Smelt)~Drought, data = FMWT)
summary(l1)
plot(l1)
emmeans(l1, pairwise ~ Drought)
visreg(l1)
```

Diagnostic plots look good on that one

Now let's try the index
```{r}
l2 = lm(log(Longfin.Smelt)~Index, data = FMWT)
summary(l2)
plot(l2)
visreg(l2)
```


Most real statistitions tell you not to log-transform count data. So usually Poisson or Negative binomial are th eway to go. 

```{r}
p1 = glm(Longfin.Smelt ~ Drought, data = FMWT, family = "poisson")
summary(p1)
plot(p1)
#that's a big bucket of NOPE
```


Now the negative binomial. We'll try a year effect in there too. 

```{r}
#ok, now how about a negative binomial?

nb1 = glm.nb(Longfin.Smelt ~ Drought, data = FMWT, link = "log")
summary(nb1)
plot(nb1)
#similar to log-transformed version, but diagnostic plots don't look as good.

nb2 = glm.nb(Longfin.Smelt ~ Index, data = FMWT)
summary(nb2)

nb3 = glmer.nb(Longfin.Smelt ~ Index + (1|Year), data = FMWT)
summary(nb3)
#not sure that's right

nb4 = glm.nb(Longfin.Smelt ~ Index + Year, data = FMWT)
summary(nb4)
plot(nb4)
visreg(nb4)

```

Well, the q-q plot isn't great, we may want to look at some alternative distributions or go to a GAM, but the basic story is pretty clear. Longfin Smelt like wet years. 