---
title: "Long-term WQ: Sampling Effort"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Look at the sampling effort for the various WQ parameters to be included in the long-term WQ publication. Here are some notes and issues to consider:

Notable differences between `raw_chla_1975_2021` (from `DroughtData`) and the data used by the Primary Producers group:

1) `raw_chla_1975_2021` includes the four EZ stations from EMP
2) `raw_chla_1975_2021` has data for November 2021
3) `raw_chla_1975_2021` has additional Chlorophyll_Sign column that indicates whether the chlorophyll-a value is above or below the RL. For the 3 values that are below the RL, `raw_chla_1975_2021` reports them as equal to the RL, while the data from the Primary Producers group reports them as half the RL.

Things to consider with the nutrient, chlorophyll-a, and WQ data sets:

1) The chlorophyll-a data set includes the Grant Line Canal and Old River subregion. Should this region be included in the nutrient and WQ data sets?
2) The chlorophyll-a data set is not filtered to remove records where `Long_term == FALSE`, while those records are removed from the nutrient and WQ data sets. We may want to make these all consistent.

So far, these are the surveys included in this summary:

* **Water Temperature, Salinity, Secchi depth:** EMP, STN, FMWT
* **Nutrients:** EMP, USGS_SFBS, USGS_CAWSC (just 11447650 - Sacramento River at Freeport)
* **Chlorophyll-a:** EMP, USGS_SFBS

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(hms)
library(scales)
library(DroughtData)
library(discretewq)
library(deltamapr)
library(sf)
library(dtplyr)
```

```{r session info}
# Run session info to display package versions
devtools::session_info()
```


# Import and Prepare Data

```{r prepare data}
# Define factor level for Region
region_lev <- c(
  "North",
  "SouthCentral",
  "Confluence",
  "Suisun Bay",
  "Suisun Marsh"
)

# Create a nested data frame and prepare to run the summary plot function on - data in DroughtData
ndf_lt_wq <- 
  tibble(
    Parameter = c(
      "Temperature",
      "Salinity",
      "Secchi",
      "DissAmmonia",
      "DissNitrateNitrite",
      "DissOrthophos",
      "Chlorophyll"
    ),
    df_data = c(
      rep(list(raw_wq_1975_2021), 3),
      rep(list(raw_nutr_1975_2021), 3),
      list(raw_chla_1975_2021)
    )
  ) %>% 
  mutate(
    df_data_c = map2(
      df_data, 
      Parameter,
      ~ drop_na(.x, all_of(.y)) %>% 
        mutate(
          Month = fct_rev(month(Date, label = TRUE)),
          Region = factor(Region, levels = region_lev)
        )
    )
  )

# Pull in and prepare the WQ data from discretewq
df_dwq <-
  wq(
    Sources = c(
      "EMP",
      "STN",
      "FMWT",
      "EDSM",
      "DJFMP",
      "SDO",
      "SKT",
      "SLS",
      "20mm",
      "Suisun",
      "Baystudy",
      "USBR",
      "USGS_SFBS",
      "YBFMP",
      "USGS_CAWSC"
    )
  ) %>% 
  transmute(
    Source,
    Station,
    Latitude,
    Longitude,
    Date = date(Date),
    # Convert Datetime to PST
    Datetime = with_tz(Datetime, tzone = "Etc/GMT+8"),
    Temperature,
    Salinity,
    Secchi,
    Chlorophyll,
    DissAmmonia,
    DissNitrateNitrite,
    DissOrthophos
  ) %>% 
  filter(
    !if_all(
      c(Temperature, Salinity, Secchi, Chlorophyll, DissAmmonia, DissNitrateNitrite, DissOrthophos),
      is.na
    )
  )
```

# Data in `DroughtData`

## Create Plots

```{r plot func}
# Create function for sampling effort plots
plot_samp_effort <- function(df, grp_var, yr_var) {
  df %>% 
    count(.data[[yr_var]], Month, .data[[grp_var]], name = "num_samples") %>% 
    ggplot(aes(x = .data[[yr_var]], y = Month, fill = num_samples)) +
    geom_tile() +
    facet_grid(rows = vars(.data[[grp_var]]), drop = FALSE) +
    scale_x_continuous(breaks = breaks_pretty(20), expand = expansion()) +
    scale_y_discrete(drop = FALSE) +
    scale_fill_viridis_c(name = "Number of Samples") +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      legend.position = "top"
    )
}
```

```{r create plots}
# Create plots for each Parameter
ndf_lt_wq_plt <- ndf_lt_wq %>% 
  mutate(plt = map(df_data_c, .f = plot_samp_effort, grp_var = "Region", yr_var = "YearAdj"))
```

## Sampling Effort {.tabset .tabset-pills}

```{r print plots, echo = FALSE, results = "asis", fig.width = 8.5, fig.height = 9}
for (i in 1:nrow(ndf_lt_wq_plt)) {
  # Create subheadings for each Parameter
  cat("### ", ndf_lt_wq_plt$Parameter[i], "\n\n")
  # Print plot
  print(ndf_lt_wq_plt$plt[[i]])
  cat("\n\n")
}
```

## Secchi Sampling Effort for North region

The secchi depth data has a large hole from Jan-Aug in 2017-2021 in the North region, which seems suspicious. Let's look at the sampling effort for secchi for each station in the North region.

```{r plot secchi sampling effort north, fig.width = 8.5, fig.height = 7}
ndf_lt_wq$df_data_c[[3]] %>% 
  filter(Region == "North") %>%
  plot_samp_effort(grp_var = "Station", yr_var = "YearAdj")
```

Hmmm, the problem seems to be coming from the one current EMP station (C3A) in the North region. Let's look at the data directly from the `discretewq` package (version 2.3.2) to see if other current EMP stations have WQ data during the same time period.

## WQ Sampling Effort for current EMP stations {.tabset .tabset-pills}

```{r wq sampling effort all EMP, message = FALSE}
# Import EMP station metadata from most recent EDI publication to use for status designation
df_emp_meta <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.458.6&entityid=ecf241d54a8335a49f8dfc8813d75609")

# Create vector of stations with active status
emp_sta_active <- df_emp_meta %>% 
  filter(Status == "Active") %>% 
  pull(Station)

# Pull out water quality data from EMP dataset within discretewq
emp_wq <- discretewq::EMP %>% 
  transmute(
    Station,
    Date = date(Date),
    Year = year(Date),
    Month = fct_rev(month(Date, label = TRUE)),
    Conductivity,
    Temperature,
    Secchi
  ) %>% 
  # Only include active stations
  filter(Station %in% emp_sta_active)

# Create a nested data frame and run the summary plot function on each WQ parameter
ndf_emp_wq <- 
  tibble(
    Parameter = c(
      "Conductivity",
      "Temperature",
      "Secchi"
    ),
    df_data = rep(list(emp_wq), 3)
  ) %>% 
  mutate(
    df_data_c = map2(df_data, Parameter, ~ drop_na(.x, all_of(.y))),
    plt = map(df_data_c, .f = plot_samp_effort, grp_var = "Station", yr_var = "Year")
  )
```

It looks like only the shore-based stations (C3A, C10A, C9) have the issue of missing secchi depth data in recent years. Sarah Perry confirmed that EMP doesn't collect secchi depth data at their shore-based stations anymore, so that explains what's going on here.

```{r print emp plots, echo = FALSE, results = "asis", fig.width = 8.5, fig.height = 17}
for (i in 1:nrow(ndf_emp_wq)) {
  # Create subheadings for each Parameter
  cat("### ", ndf_emp_wq$Parameter[i], "\n\n")
  # Print plot
  print(ndf_emp_wq$plt[[i]])
  cat("\n\n")
}
```


# Data in `discretewq`

Now that we've been making updates to the WQ data in the [`discretewq` package](https://github.com/sbashevkin/discretewq), let's take a look at which surveys we can use for the long-term WQ publication. First, we'll look at the temporal scale of all of the surveys available.

```{r discretewq temporal all}
df_dwq %>% 
  group_by(Source) %>% 
  summarize(min_date = min(Date), max_date = max(Date)) %>% 
  arrange(min_date)
```

Overall, for all parameters, it looks like FMWT, USGS_SFBS, STN, USGS_CAWSC, EMP, DJFMP, Suisun, and Baystudy have adequate temporal coverage for the long-term analysis. Now let's take a look at the spatial coverage.

```{r discretewq spatial long term}
# Only include surveys with adequate temporal coverage
df_dwq_lt <- df_dwq %>% 
  filter(Source %in% c("FMWT", "USGS_SFBS", "STN", "USGS_CAWSC", "EMP", "DJFMP", "Suisun", "Baystudy"))

# Pull out station coordinates and convert to sf object
sf_dwq_lt <- df_dwq_lt %>% 
  distinct(Source, Station, Latitude, Longitude) %>% 
  drop_na(Latitude, Longitude) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326) %>% 
  st_transform(st_crs(R_EDSM_Subregions_Mahardja_FLOAT))

# Load Delta Shapefile from Brian and only keep SubRegions east of Carquinez Straight
sf_delta <- R_EDSM_Subregions_Mahardja_FLOAT %>% 
  filter(
    !SubRegion %in% c(
      "Carquinez Strait", 
      "Lower Napa River", 
      "San Francisco Bay",
      "San Pablo Bay",
      "South Bay",
      "Upper Napa River" 
    )
  ) %>% 
  select(SubRegion)

# Plot all stations over the SubRegions
ggplot() +
  geom_sf(data = sf_delta, fill = "green", alpha = 0.5) +
  geom_sf(data = sf_dwq_lt)

# Assign SubRegions to the stations
sf_dwq_lt_reg <- sf_dwq_lt %>% 
  st_join(sf_delta, join = st_intersects) %>%
  filter(!is.na(SubRegion))

# Remove SubRegions without stations from sf_delta
sf_delta_c <- sf_delta %>% filter(SubRegion %in% unique(sf_dwq_lt_reg$SubRegion))

# Plot all stations over the SubRegions again - with limited ranges
ggplot() +
  geom_sf(data = sf_delta_c, fill = "blue", alpha = 0.5) +
  geom_sf(data = sf_dwq_lt_reg)
```

All but one of the SubRegions east of Carquinez Straight has at least one station from a long term survey within it. Now let's take a closer look at the temporal data coverage for each Station and parameter.

```{r prepare discretewq long term}
# Prepare long-term data from discretewq
df_dwq_lt_c <- df_dwq_lt %>% 
  # Remove records without latitude-longitude coordinates
  drop_na(Latitude, Longitude) %>%
  # Convert to sf object
  st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE) %>%
  # Change to crs of sf_delta_c
  st_transform(crs = st_crs(sf_delta_c)) %>%
  # Add subregions
  st_join(sf_delta_c, join = st_intersects) %>%
  # Remove any data outside our subregions of interest
  filter(!is.na(SubRegion)) %>%
  # Drop sf geometry column since it's no longer needed
  st_drop_geometry() %>% 
  # Add variables for adjusted calendar year, month, and season
    # Adjusted calendar year: December-November, with December of the previous calendar year
    # included with the following year
  mutate(
    Month = month(Date),
    YearAdj = if_else(Month == 12, year(Date) + 1, year(Date)),
    Season = case_when(
      Month %in% 3:5 ~ "Spring",
      Month %in% 6:8 ~ "Summer",
      Month %in% 9:11 ~ "Fall",
      Month %in% c(12, 1, 2) ~ "Winter"
    )
  ) %>% 
  # Restrict data to 1975-2021
  filter(YearAdj %in% 1975:2021)
```

## Sampling Effort by Station {.tabset .tabset-pills}

```{r plot func long term by sta}
plot_samp_effort_sta <- function(df, param) {
  df_param <- df %>% drop_na(.data[[param]])
  
  # Look for any instances when more than 1 data point was collected at a station-day
  df_dups <- df_param %>%
    count(Station, Date) %>% 
    filter(n > 1) %>% 
    select(-n)
  
  # Fix duplicates if there are any
  if (length(df_dups > 0)) {
    df_dups_fixed <- df_param %>%
      inner_join(df_dups, by = c("Station", "Date")) %>%
      drop_na(Datetime) %>%
      mutate(
        # Create variable for time
        Time = as_hms(Datetime),
        # Calculate difference from noon for each data point for later filtering
        Noon_diff = abs(hms(hours = 12) - Time)
      ) %>%
      # Use dtplyr to speed up operations
      lazy_dt() %>%
      group_by(Station, Date) %>%
      # Select only 1 data point per station and date, choose data closest to noon
      filter(Noon_diff == min(Noon_diff)) %>%
      # When points are equidistant from noon, select earlier point
      filter(Time == min(Time)) %>%
      ungroup() %>%
      # End dtplyr operation
      as_tibble() %>%
      select(-c(Time, Noon_diff))
    
    df_param <- df_param %>% 
      anti_join(df_dups, by = c("Station", "Date")) %>%
      bind_rows(df_dups_fixed)
  }
  
  # Create summary heat map of sampling effort for each station
  df_param %>%
    count(Station, YearAdj, name = "num_samples") %>%
    ggplot(aes(x = YearAdj, y = Station, fill = num_samples)) +
    geom_tile() +
    scale_x_continuous(breaks = breaks_pretty(20), expand = expansion()) +
    scale_fill_viridis_c(name = "Number of Samples") +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
      legend.position = "top"
    )
}
```

```{r create plots long term by sta}
# Nest long-term data from discretewq by Source
ndf_dwq_lt_c <- df_dwq_lt_c %>% nest(df_data = -Source)

# Create a nested data frame to run the summary plot function on
ndf_dwq_lt_c <- 
  tibble(
    Parameter = c(
      "Temperature",
      "Salinity",
      "Secchi",
      "DissAmmonia",
      "DissNitrateNitrite",
      "DissOrthophos",
      "Chlorophyll"
    ),
    df_data = rep(list(ndf_dwq_lt_c), 7)
  ) %>% 
  unnest(df_data)

# Create plots for each Parameter and Source
ndf_dwq_lt_plt <- ndf_dwq_lt_c %>% 
  mutate(plt = map2(df_data, Parameter, .f = plot_samp_effort_sta)) %>% 
  select(-df_data) %>% 
  nest(plts = -Parameter)
```

```{r print plots long term by sta, echo = FALSE, results = "asis", fig.width = 8, fig.height = 10}
for (i in 1:nrow(ndf_dwq_lt_plt)) {
  # Create subheadings for each Parameter
  cat("### ", ndf_dwq_lt_plt$Parameter[i], " {.tabset .tabset-pills}\n\n")
  for (j in 1:nrow(ndf_dwq_lt_plt$plts[[i]])) {
    # Create subheadings for each Survey
    cat("#### ", ndf_dwq_lt_plt$plts[[i]]$Source[j], "\n\n")
    # Print plot
    print(ndf_dwq_lt_plt$plts[[i]]$plt[[j]])
    cat("\n\n")
  }
}
```

