---
title: "EDB: Continuous Chlorophyll Figures"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Create some time-series and boxplot figures of the daily averages and medians of the continuous chlorophyll data for 2020 and 2021 for the **February 2022** Emergency Drought Barrier report.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
library(rlang)
library(patchwork)
# Install the EDBdata package version 0.2.0 to use the data displayed in the
  # figures of the February 2022 version of the EDB report. Update EDBdata to >=
  # v0.2.1 to update the figures for a future version of the report. This will
  # update the daily averages and medians of the two USGS stations - MDM and SJJ.
# install.packages("devtools")
# devtools::install_github("mountaindboz/EDBdata", ref = "v0.2.0")
library(EDBdata)
```

# Import and Prepare Data

```{r import data, message = FALSE}
# Import discrete chlorophyll data for OSJ, FAL, and D19
df_d_chla_orig <- read_csv("EDB/d_chla_fal_osj_d19.csv")
```

```{r prepare data}
# Prepare data for all figures
df_cont_chla_clean <- cont_chla_daily  %>% 
  # Only include Apr-Dec
  filter(month(Date) %in% 4:12) %>% 
  # Limit the stations included in the figures
  filter(!Station %in% c("ORI", "WCI")) %>% 
  # Reassign FAL to the Central Delta region
  mutate(Region = if_else(Station == "FAL", "Central Delta", Region)) %>% 
  # Assign factor order to Region
  mutate(Region = factor(Region, levels = c("Sacramento", "San Joaquin", "Central Delta")))
  
# Continue to prepare data for the time-series figures
df_cont_chla_ts <- df_cont_chla_clean %>% 
  # Fill in missing dates with NA values for geom_line to not interpolate data gaps
  group_by(Station, Region, Year) %>% 
  complete(Date = seq.Date(min(Date), max(Date), by = "1 day")) %>%
  ungroup() 

# Continue to prepare data for the boxplot figures
df_cont_chla_bp <- df_cont_chla_clean %>% 
  # Add a variable for Season
  mutate(
    Month = month(Date, label = TRUE),
    Season = factor(
      case_when(
        Month %in% c("Apr", "May", "Jun") ~ "Spring",
        Month %in% c("Jul", "Aug", "Sep") ~ "Summer",
        Month %in% c("Oct", "Nov", "Dec") ~ "Fall"
      ),
      levels = c("Spring", "Summer", "Fall")
    )
  )
  
# Convert data into nested dataframe to apply time-series plotting function by Region
ndf_cont_chla_ts <- df_cont_chla_ts %>% 
  nest(df_data = -Region) %>% 
  arrange(Region)

# Create a tibble of the EDB time periods to add shaded rectangles to the time-series plots
df_edb_dates <- 
  tibble(
    Year = 2020:2021,
    EDB_const_st = ymd(c(NA_Date_, "2021-06-03")),
    EDB_const_end = ymd(c(NA_Date_, "2021-06-23")),
    EDB_end = ymd(c(NA_Date_, "2022-01-01"))
  )

# Create a tibble to define consistent min and max values for the y-axes of the time-series plots
df_ts_ylim <- df_cont_chla_clean %>% 
  summarize(across(starts_with("Chla_"), list(min = ~ floor(min(.x)), max = ~ ceiling(max(.x)))))

# Prepare 2021 discrete chlorophyll data for FAL, OSJ, and D19 to be added to time-series plots
  # of continuous data
df_d_chla_clean <- df_d_chla_orig %>% 
  filter(Date >= "2021-04-01") %>% 
  mutate(Station = if_else(Station == "D19", "FRK", Station))
```

# Create Figures

```{r plot functions}
# Create a list for custom formatting of EDB plots
edb_theme <- list(
  theme_light(),
  theme(
    strip.text = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )
)

# Function for time-series plots of continuous chlorophyll data by Region
ts_plot_region <- function(df, region_var, data_var) {
  
  # Define color palettes based on region designation
  color_pal <- switch (as.character(region_var),
    "San Joaquin" = scale_color_viridis_d(option = "plasma", end = 0.8),
    "Central Delta" = scale_color_viridis_d(option = "plasma", end = 0.8),
    Sacramento = scale_color_viridis_d(option = "viridis", end = 0.8)
  )
  
  # Convert data_var to a string to be used in conditionals
  data_var_str <- as_name(ensym(data_var))
  
  # Define min and max limits for the y-axes based on data_var
  y_min <- if (data_var_str == "Chla_avg") df_ts_ylim$Chla_avg_min else df_ts_ylim$Chla_med_min
  y_max <- if (data_var_str == "Chla_avg") df_ts_ylim$Chla_avg_max else df_ts_ylim$Chla_med_max
  
  # Define y-axis label
  if (region_var == "San Joaquin") {
    if (data_var_str == "Chla_avg") {
      y_label <- "Daily Average Chlorophyll (µg/L)"
    } else {
      y_label <- "Daily Median Chlorophyll (µg/L)"
    }
  } else {
    y_label <- NULL
  }
  
  # Define x-axis label
  x_label <- if (region_var == "Central Delta") "Date" else NULL
  
  # Create plot
  p <- df %>% 
    ggplot(aes(x = Date, y = {{ data_var }}, color = Station)) +
    geom_line() +
    facet_wrap(vars(Year), nrow = 1, scales = "free_x") +
    ggtitle(region_var) +
    edb_theme +
    # Add shaded rectangles for EDB time periods
    geom_rect(
      aes(xmin = EDB_const_st, xmax = EDB_const_end, ymin = -Inf, ymax = Inf),
      data = df_edb_dates,
      inherit.aes = FALSE,
      alpha = 0.2,
      fill = "steelblue"
    ) +
    geom_rect(
      aes(xmin = EDB_const_end, xmax = EDB_end, ymin = -Inf, ymax = Inf),
      data = df_edb_dates,
      inherit.aes = FALSE,
      alpha = 0.2,
      fill = "grey30"
    ) +
    color_pal +
    scale_y_continuous(name = y_label, limits = c(y_min, y_max)) +
    scale_x_date(
      name = x_label,
      breaks = breaks_pretty(10),
      labels = label_date_short(format = c(NA, "%b", NA, NA)),
      expand = expansion(mult = 0.035)
    )
  
  return(p)
}
```

```{r create ts figures}
# Daily averages
plt_ts_avg <- ndf_cont_chla_ts %>% 
  mutate(plt_ts = map2(df_data, Region, .f = ts_plot_region, data_var = Chla_avg)) %>% 
  pull(plt_ts) %>% 
  wrap_plots(ncol = 1)

# Daily medians
plt_ts_med <- ndf_cont_chla_ts %>% 
  mutate(plt_ts = map2(df_data, Region, .f = ts_plot_region, data_var = Chla_med)) %>% 
  pull(plt_ts) %>% 
  wrap_plots(ncol = 1)
```

## Time-series Figures

### Daily Averages

The light blue shading represents when barrier was under construction, and the grey shading represents when the barrier was complete and in place.

```{r print daily avg ts figure, fig.height = 10, fig.width = 8, warning = FALSE}
plt_ts_avg
```

### Daily Medians

The light blue shading represents when barrier was under construction, and the grey shading represents when the barrier was complete and in place.

```{r print daily med ts figure, fig.height = 10, fig.width = 8, warning = FALSE}
plt_ts_med
```

### Discrete and Continuous Comparison

The light blue shading represents when barrier was under construction, and the grey shading represents when the barrier was complete and in place.

```{r ts figure disc and cont comp, fig.height = 6, fig.width = 5, warning = FALSE}
plt_ts_d_cont <- df_cont_chla_ts %>% 
  # Only include FAL, OSJ, and FRK for 2021
  filter(
    Year == 2021,
    Station %in% c("FAL", "OSJ", "FRK")
  ) %>% 
  ggplot(aes(x = Date, y = Chla_med)) +
  geom_line() +
  facet_wrap(
    vars(Station), 
    ncol = 1,
    labeller = labeller(Station = c("FAL" = "FAL", "OSJ" = "OSJ", "FRK" = "FRK (D19)"))
  ) +
  edb_theme +
  scale_x_date(
    breaks = breaks_pretty(10),
    labels = label_date_short(),
    expand = expansion(mult = 0.02)
  ) +
  ylab("Chlorophyll (µg/L)") +
  # Add discrete chlorophyll points
  geom_point(
    data = df_d_chla_clean,
    aes(x = Date, y = Chla),
    shape = 21,
    size = 2,
    color = "grey30",
    fill = "green"
  ) +
  # Add shaded rectangles for EDB time periods
  geom_rect(
    aes(xmin = EDB_const_st, xmax = EDB_const_end, ymin = -Inf, ymax = Inf),
    data = df_edb_dates,
    inherit.aes = FALSE,
    alpha = 0.2,
    fill = "steelblue"
  ) +
  geom_rect(
    aes(xmin = EDB_const_end, xmax = EDB_end, ymin = -Inf, ymax = Inf),
    data = df_edb_dates,
    inherit.aes = FALSE,
    alpha = 0.2,
    fill = "grey30"
  ) 

plt_ts_d_cont
```

## Boxplot Figures

### By Season

```{r boxplot by season, fig.width = 8}
plt_bp_seas <- df_cont_chla_bp %>% 
  ggplot(aes(x = Region, y = Chla_med, fill = Season)) +
  geom_boxplot() +
  facet_wrap(vars(Year)) +
  edb_theme +
  ylab("Daily Median Chlorophyll (µg/L)") +
  scale_fill_viridis_d(option = "plasma")

plt_bp_seas
```

### By Year

```{r boxplot by year}
plt_bp_yr <- df_cont_chla_bp %>% 
  mutate(Year = as.character(Year)) %>% 
  ggplot(aes(x = Region, y = Chla_med, fill = Year)) +
  geom_boxplot() +
  edb_theme +
  ylab("Daily Median Chlorophyll (µg/L)") +
  scale_fill_viridis_d(end = 0.8)

plt_bp_yr
```

# Export Figures for Report

All figures for the February 2022 report exported here were moved to the EDB SharePoint site.

```{r export rpt figures, eval = FALSE}
# Export daily average time-series figure
ggsave(
  "EDB/cont_chla_ts_daily_avg.jpg",
  plot = plt_ts_avg,
  width = 6.5,
  height = 8.25,
  units = "in"
)

# Export daily median time-series figure
ggsave(
  "EDB/cont_chla_ts_daily_med.jpg",
  plot = plt_ts_med,
  width = 6.5,
  height = 8.25,
  units = "in"
)

# Export discrete and continuous time-series comparison figure
ggsave(
  "EDB/cont_d_chla_ts_comp.jpg",
  plot = plt_ts_d_cont,
  width = 4.5,
  height = 5,
  units = "in"
)

# Export boxplot by season figure
ggsave(
  "EDB/cont_chla_bp_season.jpg",
  plot = plt_bp_seas,
  width = 6.5,
  height = 4,
  units = "in"
)

# Export boxplot by year figure
ggsave(
  "EDB/cont_chla_bp_yr.jpg",
  plot = plt_bp_yr,
  width = 5,
  height = 4,
  units = "in"
)
```

