---
title: "EDB Phyto Data Overview"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "-")
```

# Purpose

Preliminary data exploration of the phytoplankton data for 2014 - 2021 for the Emergency Drought Barrier report.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
# install.packages("devtools")
# devtools::install_github("mountaindboz/EDBdata")
library(EDBdata)
```

# Sampling Effort Counts

## Stations and Years

```{r count stations years}
# See if the sampling effort changed at stations between years
phyto_edb %>%
  count(Year, Region, Station) %>%
  arrange(Year) %>%
  pivot_wider(names_from = Year, values_from = n) %>%
  arrange(Region, Station) %>% 
  kable()
```

## Months and Years

```{r count months years}
# See if the sampling effort changed between months and years
phyto_edb %>%
  mutate(Month = month(DateTime, label = TRUE)) %>%
  count(Year, Month) %>%
  arrange(Month) %>%
  pivot_wider(names_from = Month, values_from = n) %>%
  arrange(Year) %>% 
  kable()
```

## Stations, Months, and Years

```{r count stations months years}
# Overall view of sampling effort - stations, months, and years
phyto_edb %>%
  mutate(Month = month(DateTime, label = TRUE)) %>%
  count(Year, Month, Region, Station) %>%
  arrange(Month) %>%
  pivot_wider(names_from = Month, values_from = n) %>%
  complete(Year, nesting(Region, Station)) %>% 
  arrange(Region, Station, Year) %>% 
  kable()
```

# Exploratory Plots

```{r general preparation}
# Prepare phyto data for all plots
phyto_edb_c <- phyto_edb %>% 
  mutate(
    Year = as.character(Year),
    Region = factor(Region, levels = c("Sacramento", "San Joaquin", "Central Delta"))
  )
```

## Total Density

Calculate total organisms per mL for each station and sampling date

```{r prepare total density, message = FALSE}
# Calculate total density of organisms for each station and sampling date
phyto_total_all <- phyto_edb_c %>% 
  group_by(Station, Region, Year, Date) %>% 
  summarize(TotalDens = sum(OrganismsPerMl)) %>% 
  ungroup()

# Create a list of common ggplot elements for the boxplots
lst_phyto_boxplt <- list(
  scale_y_continuous(
    name = "Total Organisms per mL", 
    labels = label_comma()
  ),
  theme_bw()
)
```

### Boxplots by Year

```{r boxplot total density by year}
phyto_total_all %>% 
  ggplot(aes(x = Year, y = TotalDens)) +
  geom_boxplot() +
  lst_phyto_boxplt
```

### Boxplots by Region

Here are the region designations to better interpret the jitter points in the plot:

* Sacramento: D22, D24, D4, NZ068
* San Joaquin: D12, D16, D26
* Central Delta: C9, D19, D28A

```{r boxplot total density by region}
phyto_total_all %>%
  ggplot(aes(x = Region, y = TotalDens)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = Station), width = 0.3) +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.95)
```

### Boxplots by Year and Region

```{r boxplot total density by year region}
phyto_total_all %>% 
  ggplot(aes(x = Year, y = TotalDens, color = Region)) +
  geom_boxplot() +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  theme(legend.position = "top")
```

2014, 2015, 2020, and 2021 are difficult to see. Let's look at just those.

```{r boxplot total density by year region dry years}
phyto_total_all %>% 
  filter(Year %in% c("2014", "2015", "2020", "2021")) %>% 
  ggplot(aes(x = Year, y = TotalDens, color = Region)) +
  geom_boxplot() +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9)
```

Let's break these boxplots apart by season. Since we are interested in the impact of the Barrier, we should consider that in how we define our seasons.

* Both 2021 and 2015 barriers were in place June - November
* 2021 barrier was also in place in December
* Notch in 2021 barrier January - April 2022
* Maybe we'll use the following for our season definitions:
	* Summer: June - August
	* Fall: September - December
	* Spring: January - May

```{r add seasons}
phyto_total_all_seas <- phyto_total_all %>% 
  mutate(
    Month = month(Date),
    Season = factor(
      case_when(
        Month %in% c(1:5) ~ "Spring",
        Month %in% c(6:8) ~ "Summer",
        Month %in% c(9:12) ~ "Fall"
      ),
      levels = c("Spring", "Summer", "Fall")
    )
  )
```

### Boxplots by Season

```{r boxplot total density by season}
phyto_total_all_seas %>% 
  ggplot(aes(x = Season, y = TotalDens)) +
  geom_boxplot() +
  lst_phyto_boxplt
```

### Boxplots by Year and Season

```{r boxplot total density by year season}
phyto_total_all_seas %>% 
  ggplot(aes(x = Year, y = TotalDens, color = Season)) +
  geom_boxplot() +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  theme(legend.position = "top")
```

Again, 2014, 2015, 2020, and 2021 are difficult to see. Let's look at just those.

```{r boxplot total density by year season dry years}
phyto_total_all_seas %>% 
  filter(Year %in% c("2014", "2015", "2020", "2021")) %>%
  ggplot(aes(x = Year, y = TotalDens, color = Season)) +
  geom_boxplot() +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9)
```

### Boxplots by Region and Season

```{r boxplot total density by region season}
phyto_total_all_seas %>% 
  ggplot(aes(x = Region, y = TotalDens, color = Season)) +
  geom_boxplot() +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9)
```

### Boxplots by Year, Region, and Season

```{r boxplot total density by year region season, fig.height = 9, fig.width = 8}
phyto_total_all_seas %>% 
  ggplot(aes(x = Season, y = TotalDens, color = Region)) +
  geom_boxplot() +
  facet_wrap(vars(Year)) +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  theme(legend.position = "top")
```

Again, let's look at just 2014, 2015, 2020, and 2021.

```{r boxplot total density by year region season dry years}
phyto_total_all_seas %>% 
  filter(Year %in% c("2014", "2015", "2020", "2021")) %>%
  ggplot(aes(x = Season, y = TotalDens, color = Region)) +
  geom_boxplot() +
  facet_wrap(vars(Year)) +
  lst_phyto_boxplt +
  scale_color_viridis_d(option = "plasma", end = 0.9) +
  theme(legend.position = "top")
```

## Relative Density by Algal Type

Calculate total organisms per mL for each station, sampling date, and algal type

```{r prepare total density by algal type, message = FALSE}
# Calculate total density of organisms for each station, sampling date, and algal type
phyto_total_algalt <- phyto_edb_c %>% 
  mutate(
    Month = month(Date),
    Season = factor(
      case_when(
        Month %in% c(1:5) ~ "Spring",
        Month %in% c(6:8) ~ "Summer",
        Month %in% c(9:12) ~ "Fall"
      ),
      levels = c("Spring", "Summer", "Fall")
    ),
    Month = month(Date, label = TRUE)
  ) %>% 
  group_by(Station, Region, Year, Season, Month, Date, AlgalType) %>% 
  summarize(TotalDens = sum(OrganismsPerMl)) %>% 
  ungroup()

# Create a list of common ggplot elements for the relative density barplots
lst_phyto_rd_barplt <- list(
  scale_y_continuous(
    name = "Relative Density",
    labels = label_percent()
  ),
  scale_fill_discrete(name = "Algal Type"),
  theme_bw()
)
```

### Barplot by Year

```{r barplot rel density by year, message = FALSE}
phyto_total_algalt %>%
  group_by(Year, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Year, y = AvgDens, fill = AlgalType)) +
  geom_col(position = "fill") +
  lst_phyto_rd_barplt
```

### Barplot by Region

```{r barplot rel density by region, message = FALSE}
phyto_total_algalt %>%
  group_by(Region, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Region, y = AvgDens, fill = AlgalType)) +
  geom_col(position = "fill") +
  lst_phyto_rd_barplt
```

### Barplot by Season

```{r barplot rel density by season, message = FALSE}
phyto_total_algalt %>%
  group_by(Season, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Season, y = AvgDens, fill = AlgalType)) +
  geom_col(position = "fill") +
  lst_phyto_rd_barplt
```

### Barplot by Month

```{r barplot rel density by month, message = FALSE}
phyto_total_algalt %>%
  group_by(Month, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Month, y = AvgDens, fill = AlgalType)) +
  geom_col(position = "fill") +
  lst_phyto_rd_barplt
```

Cyanobacteria dominate all grouping variables. Let's see what happens when we remove this Algal Type.

```{r remove cyano from total density}
phyto_total_algalt_nocyan <- phyto_total_algalt %>% filter(AlgalType != "Cyanobacterium")
```

### Barplot by Year and Region

```{r barplot rel density by year region no cyano, message = FALSE, fig.height = 6}
phyto_total_algalt_nocyan %>%
  group_by(Year, Region, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Region, y = AvgDens, fill = AlgalType)) +
  geom_col(color = "grey50", position = "fill") +
  facet_wrap(vars(Year)) +
  lst_phyto_rd_barplt +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

### Barplot by Year and Season

```{r barplot rel density by year season no cyano, message = FALSE}
phyto_total_algalt_nocyan %>%
  group_by(Year, Season, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Season, y = AvgDens, fill = AlgalType)) +
  geom_col(color = "grey50", position = "fill") +
  facet_wrap(vars(Year)) +
  lst_phyto_rd_barplt
```

### Barplot by Year, Region, and Season

```{r barplot rel density by year region season no cyano, message = FALSE, fig.height = 9, fig.width = 7}
phyto_total_algalt_nocyan %>%
  group_by(Year, Region, Season, AlgalType) %>%
  summarize(AvgDens = mean(TotalDens)) %>%
  ungroup() %>%
  ggplot(aes(x = Season, y = AvgDens, fill = AlgalType)) +
  geom_col(color = "grey50", position = "fill") +
  facet_grid(
    cols = vars(Region),
    rows = vars(Year)
  ) +
  lst_phyto_rd_barplt +
  theme(legend.position = "top")
```

