---
title: "HAB Satellite Data - Exploration and Analysis"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

Data exploration and analysis of the HAB satellite data for Franks Tract, Mildred Island, Clifton Court Forebay, and Liberty Island in 2020-2021. Results will be presented in the **June 2022 version** of the HABs/Weeds report for the Emergency Drought Barrier.

# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(lubridate)
library(EDBdata)
library(sf)
library(deltamapr)
library(curl)
library(glue)
library(stars)
library(car)
library(knitr)
library(kableExtra)
library(patchwork)
library(here)
```

```{r check dir}
# Check if we are in the correct working directory
i_am("EDB/HAB_satellite_analysis_June2022.Rmd")
```

# Import and Prepare Data

Import the average Cyano Index data from a csv file since it may not remain in the `EDBdata` package.

```{r import avg ci data, message = FALSE}
# Import average Cyano Index data
df_hab_sat_avg_ci <- read_csv(here("EDB/hab_sat_ow_delta_avg_ci.csv")) %>% 
  mutate(Date = mdy(Date))
```

Prepare both the average Cyano Index data and the pixel count data within four Cyano Index categories for figures and analysis.

```{r prepare avg ci and count data}
# Prepare both HAB satellite data sets for figures and analysis
lst_hab_sat_clean <- lst(df_hab_sat_avg_ci, hab_sat_ow_delta) %>% 
  map(
    ~ mutate(
      .x,
      YearFct = factor(year(Date)),
      Region = factor(
        Region,
        levels = c(
          "Liberty Island",
          "Franks Tract",
          "Mildred Island",
          "Clifton Court Forebay"
        )
      ),
      DOY = yday(Date)
    ) %>% 
    relocate(YearFct, .before = Region)
  )
  
# Pull out each data frame in the list so they can be called directly
df_hab_sat_avg_ci_c <- lst_hab_sat_clean$df_hab_sat_avg_ci
df_hab_sat_counts_c <- lst_hab_sat_clean$hab_sat_ow_delta
```

# Data Exploration

## All Data

### Imagery Coverage

Generate a table of imagery counts by `YearFct` and `Region`:

```{r imagery counts all}
df_hab_sat_counts_c %>% 
  count(YearFct, Region) %>% 
  pivot_wider(names_from = YearFct, values_from = n) %>% 
  kable() %>% 
  kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE,
    position = "left"
  )
```

Imagery counts are somewhat unbalanced between groups. Let's look at the imagery coverage adding the percent of valid pixels for each image:

```{r imagery coverage all, fig.width = 8, fig.height = 6}
df_hab_sat_counts_c %>%
  mutate(
    TotalPixels = rowSums(across(where(is.integer))),
    PercValid = (1 - InvalidOrMissing / TotalPixels) * 100,
    Month = fct_rev(month(Date, label = TRUE)),
    Day = day(Date)
  ) %>% 
  ggplot(aes(x = Day, y = Month)) +
  geom_tile(aes(fill = PercValid)) +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct)
  ) +
  scale_fill_viridis_c(
    name = "Percent\nValid\nPixels", 
    direction = -1
  ) +
  scale_x_continuous(
    name = "Day of Month",
    breaks = seq.int(1, 31, by = 2),
    expand = expansion()
  ) +
  scale_y_discrete(expand = expansion()) +
  theme_bw()
```

### Average Cyano Index

Generate a table of summary statistics of average Cyano Index values grouped by `YearFct` and `Region`:

```{r avg ci summary stats all, message = FALSE}
df_hab_sat_avg_ci_c %>% 
  group_by(YearFct, Region) %>% 
  summarize(
    Min = min(AvgCI),
    FirstQuantile = quantile(AvgCI, 0.25),
    Median = median(AvgCI),
    Mean = mean(AvgCI),
    ThirdQuantile = quantile(AvgCI, 0.75),
    Max = max(AvgCI),
    StdDev = sd(AvgCI)
  ) %>% 
  ungroup() %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped")
```

Let's look at time-series plots of average Cyano Index values grouped by `YearFct` and `Region`:

```{r avg ci ts plots all, fig.height = 6}
df_hab_sat_avg_ci_c %>% 
  ggplot(aes(x = Date, y = AvgCI, color = Region, fill = Region)) +
  geom_point() +
  geom_line() +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct),
    scales = "free_x"
  ) +
  ylab("Average Cyano Index") +
  scale_x_date(
    name = "Date",
    breaks = breaks_pretty(10),
    labels = label_date_short(c(NA, "%b", "%d", NA)),
    expand = expansion(mult = 0.035)
  ) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.8,
    aesthetics = c("color", "fill")
  ) +
  guides(color = "none", fill = "none") +
  theme_bw()
```

Looking at the time-series plots, it's probably best to only include June-October for each year in the analysis since we only see Cyano Index values greater than zero during these months. We'll modify the data set, and rerun the plots and tables above.

## June-October Data

```{r filter data to jun to oct}
# Restrict data to Jun-Oct for both years
df_hab_sat_avg_ci_jun_oct <- df_hab_sat_avg_ci_c %>% filter(month(Date) >= 6 & month(Date) <= 10)
df_hab_sat_counts_jun_oct <- df_hab_sat_counts_c %>% filter(month(Date) >= 6 & month(Date) <= 10)
```

### Imagery Coverage

Generate a table of imagery counts by `YearFct` and `Region`:

```{r imagery counts jun to oct}
df_hab_sat_counts_jun_oct %>% 
  count(YearFct, Region) %>% 
  pivot_wider(names_from = YearFct, values_from = n) %>% 
  kable() %>% 
  kable_styling(
    bootstrap_options = "striped",
    full_width = FALSE,
    position = "left"
  )
```

Imagery counts are still somewhat unbalanced between groups. Let's look at the imagery coverage adding the percent of valid pixels for each image:

```{r imagery coverage jun to oct, fig.width = 8, fig.height = 6}
df_hab_sat_counts_jun_oct %>%
  mutate(
    TotalPixels = rowSums(across(where(is.integer))),
    PercValid = (1 - InvalidOrMissing / TotalPixels) * 100,
    Month = fct_rev(month(Date, label = TRUE)),
    Day = day(Date)
  ) %>% 
  ggplot(aes(x = Day, y = Month)) +
  geom_tile(aes(fill = PercValid)) +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct)
  ) +
  scale_fill_viridis_c(
    name = "Percent\nValid\nPixels", 
    direction = -1
  ) +
  scale_x_continuous(
    name = "Day of Month",
    breaks = seq.int(1, 31, by = 2),
    expand = expansion()
  ) +
  scale_y_discrete(expand = expansion()) +
  theme_bw()
```

### Average Cyano Index

Generate a table of summary statistics of average Cyano Index values grouped by `YearFct` and `Region`:

```{r avg ci summary stats jun to oct, message = FALSE}
df_hab_sat_avg_ci_jun_oct %>% 
  group_by(YearFct, Region) %>% 
  summarize(
    Min = min(AvgCI),
    FirstQuantile = quantile(AvgCI, 0.25),
    Median = median(AvgCI),
    Mean = mean(AvgCI),
    ThirdQuantile = quantile(AvgCI, 0.75),
    Max = max(AvgCI),
    StdDev = sd(AvgCI)
  ) %>% 
  ungroup() %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped")
```

Let's look at time-series plots of average Cyano Index values grouped by `YearFct` and `Region`:

```{r avg ci ts plots jun to oct, fig.height = 6}
df_hab_sat_avg_ci_jun_oct %>% 
  ggplot(aes(x = Date, y = AvgCI, color = Region, fill = Region)) +
  geom_point() +
  geom_line() +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct),
    scales = "free_x"
  ) +
  ylab("Average Cyano Index") +
  scale_x_date(
    name = "Date",
    breaks = breaks_pretty(10),
    labels = label_date_short(c(NA, "%b", "%d", NA)),
    expand = expansion(mult = 0.035)
  ) +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.8,
    aesthetics = c("color", "fill")
  ) +
  guides(color = "none", fill = "none") +
  theme_bw()
```

We can also look at boxplots and histograms of the average Cyano Index values grouped by `YearFct` and `Region`:

```{r avg ci box plots jun to oct}
df_hab_sat_avg_ci_jun_oct %>% 
  ggplot(aes(x = Region, y = AvgCI, fill = Region)) +
  geom_boxplot() +
  facet_grid(cols = vars(YearFct)) +
  ylab("Average Cyano Index") +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.8
  ) +
  guides(fill = "none") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

```{r avg ci hist plots jun to oct, message = FALSE, fig.height = 6}
df_hab_sat_avg_ci_jun_oct %>% 
  ggplot(aes(x = AvgCI, fill = Region)) +
  geom_histogram() +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct)
  ) +
  scale_x_continuous(
    name = "Average Cyano Index",
    expand = expansion(mult = 0.02)
  ) +
  scale_fill_viridis_d(
    option = "plasma",
    end = 0.8
  ) +
  guides(fill = "none") +
  theme_bw()
```

We can also look at frequency polygon plots which are similar to histograms:

```{r avg ci freqpoly plots jun to oct, message = FALSE}
# All Years combined
df_hab_sat_avg_ci_jun_oct %>% 
  ggplot(aes(x = AvgCI, color = Region)) +
  geom_freqpoly() +
  scale_x_continuous(
    name = "Average Cyano Index",
    expand = expansion(mult = 0.02)
  ) +
  ggtitle("All Years Combined") +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.8
  ) +
  theme_bw()

# Separated by Year
df_hab_sat_avg_ci_jun_oct %>% 
  ggplot(aes(x = AvgCI, color = Region)) +
  geom_freqpoly() +
  facet_grid(cols = vars(YearFct)) +
  scale_x_continuous(
    name = "Average Cyano Index",
    expand = expansion(mult = 0.02)
  ) +
  ggtitle("Separated by Year") +
  scale_color_viridis_d(
    option = "plasma",
    end = 0.8
  ) +
  theme_bw() +
  theme(legend.position = "top")
```

# ANOVA analysis

This data set has a lot of zeros in it. We substituted zero for all values below the detection limit when calculating the averages, so apparently there were a lot of instances when the Cyano Index was below the detection limit. We'll try to run a linear model on this data set to look for differences in the Cyano Index between regions and years, but I am certain that this data set will not meet the model assumptions. We'll start with a linear model and go from there after taking a look at the model residuals.

## Linear Model - no transformation

```{r avg ci lm}
lm_hab_sat <- lm(
  AvgCI ~ YearFct * Region + DOY, 
  data = df_hab_sat_avg_ci_jun_oct
)

Anova(lm_hab_sat, type = 3)
```

```{r avg ci lm diag, message = FALSE}
df_hab_sat_lm_diag <- df_hab_sat_avg_ci_jun_oct %>%
  mutate(
    Residuals = residuals(lm_hab_sat),
    Fitted = predict(lm_hab_sat)
  )

# Histogram of the residuals
df_hab_sat_lm_diag %>% 
  ggplot(aes(x = Residuals)) +
  geom_histogram() +
  labs(
    title = "Residual Histogram", 
    x = "Residuals"
  ) + 
  theme_bw()

# Q-Q plot of the residuals to look at normality
df_hab_sat_lm_diag %>% 
  ggplot(aes(sample = Residuals)) + 
  labs(
    title = "Residual Probability Plot", 
    x = "Normal Quantiles", 
    y = "Residuals"
  ) + 
  stat_qq() + 
  stat_qq_line(
    size = 1, 
    color = 'red'
  ) + 
  theme_bw()

# Residuals vs. fitted values
df_hab_sat_lm_diag %>% 
  ggplot(aes(x = Fitted, y = Residuals, color = Region)) +
  geom_point() +
  labs(
    title = "Residuals vs. Fitted Values", 
    x = "Fitted Values",
    y = "Residuals"
  ) +
   scale_color_viridis_d(
    option = "plasma",
    end = 0.8
  ) +
  theme_bw() +
  theme(legend.position = "top")

# Observed vs. fitted values
df_hab_sat_lm_diag %>%
  ggplot(aes(x = Fitted, y = AvgCI)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(
    title = "Observed vs. Fitted Values", 
    x = "Fitted Values",
    y = "Observed Values"
  ) +
  theme_bw()
```

The linear model does not meet the assumptions of normally distributed residuals and constant variance across the fitted values, so we'll need to try something else. Let's try various transformations of the response variable `AvgCI`.

## Try various transformations

```{r hist function}
# Function to create histograms displaying various transformations
create_hist <- function(df, data_var, trans_label) {
  hist_all <- df %>% 
    ggplot(aes(x = {{ data_var }})) +
    geom_histogram() +
    scale_x_continuous(
      name = NULL,
      expand = expansion(mult = 0.02)
    ) +
    ggtitle("All Data") +
    theme_bw()
  
  hist_facet <- df %>% 
    ggplot(aes(x = {{ data_var }}, fill = Region)) +
    geom_histogram() +
    facet_grid(
      rows = vars(Region),
      cols = vars(YearFct)
    ) +
    scale_x_continuous(
      name = paste0("Average Cyano Index (", trans_label, ")"),
      expand = expansion(mult = 0.02)
    ) +
    scale_fill_viridis_d(
      option = "plasma",
      end = 0.8
    ) +
    guides(fill = "none") +
    ggtitle("Facetted by Region and Year") +
    theme_bw()
  
  (hist_all + plot_spacer()) / hist_facet + plot_layout(heights = c(1, 2))
}
```

Log transformation after adding 1 to `AvgCI` to account for the zeros:

```{r log avg ci, fig.height = 10, message = FALSE}
df_hab_sat_avg_ci_jun_oct %>% 
  mutate(log_AvgCI = log(AvgCI + 1)) %>%
  create_hist(log_AvgCI, "log transformed")
```

This looks almost exactly the same as the non-transformed data. This is probably because adding the one before taking the natural log swamps out the relatively small numbers in the non-transformed data.

Square Root:

```{r sqrt avg ci, fig.height = 10, message = FALSE}
df_hab_sat_avg_ci_jun_oct %>% 
  mutate(sqrt_AvgCI = sqrt(AvgCI)) %>% 
  create_hist(sqrt_AvgCI, "Square Root transformed")
```

Cube Root:

```{r cube rt avg ci, fig.height = 10, message = FALSE}
df_hab_sat_avg_ci_jun_oct %>% 
  mutate(CubeRt_AvgCI = AvgCI^(1/3)) %>% 
  create_hist(CubeRt_AvgCI, "Cube Root transformed")
```

Reciprocal Root after adding 1 to `AvgCI` to account for the zeros:

```{r recip rt avg ci, fig.height = 10, message = FALSE}
df_hab_sat_avg_ci_jun_oct %>% 
  mutate(RecipRt_AvgCI = 1 / sqrt(AvgCI + 1)) %>% 
  create_hist(RecipRt_AvgCI, "Reciprocal Root transformed")
```

Unfortunately, none of these transformations were able to make the `AvgCI` appear close to a normal distribution. There seems to be just too many zeros in the data. Other approaches to handle data with a lot of zeros such as a negative binomial model or zero-inflated model need the response variable to be count data; however, the `AvgCI` is not count data, so these approaches won't work. It seems like we may need to go back to the drawing board on this one. A couple of options to pursue:

1) Try using a `glm` with a gamma distribution. Gamma distributions require all values are greater than zero, so we may need to add 1 to the `AvgCI` to make this work. We could also substitute either the detection limit or half the detection limit instead of zero for the pixels that are below the detection limit when calculating the `AvgCI`. I'm not sure if this is appropriate though.
2) Use a `glm` with another distribution that I am not familiar with.
3) Simplify the analysis by using the relative percentages of counts of valid pixels within four Cyano Index categories (see below). Beyond visualizing the results by Region and Year, we could assign an ordinal category to each day-region combination such as Absent, Low, or High based on some predetermined thresholds for the relative percentages. We could then use this ordinal categorical variable as the response variable in an ordinal regression model. This adds another level of abstraction though so it may not be desirable.
4) Don’t try using a statistical approach and stick to only displaying the results with the area plot.

**For now, we are going to go with option #4 above. At some point, we may revisit using a statistical approach.**

# Cyano Index Categorical Analysis

Another way to visualize this data set is to look at the relative percentages of counts of valid pixels within four Cyano Index categories (Low, Moderate, High, and Very High) based on WHO recreational guidance level thresholds. We'll visualize this with area plots broken out by Region and Year and with maps of notable days in 2021.

```{r set global opts ci cat analysis}
# Define global options for the Cyano Index categorical analysis:
# Define factor levels of the Cyano Index categories
ci_cat_levels <- c(
  "NonDetect", 
  "VeryHigh",
  "High", 
  "Moderate", 
  "Low"
)

# Define factor labels of the Cyano Index categories
ci_cat_labels <- c(
  "Non Detect", 
  "Very High",
  "High", 
  "Moderate", 
  "Low"
)

# Define custom color palette for the Cyano Index categories
ci_cat_color_pal <- c(
  "Non Detect" = "#CCCCCC",
  "Low" = "#7902A8FF",
  "Moderate" = "#C5407EFF",
  "High" = "#F48849FF",
  "Very High" = "#F7E225FF"
)
```

## Area Plot

```{r prepare data area plot}
# Prepare HAB satellite data for stacked area plot
df_hab_sat_area_plt <- df_hab_sat_counts_jun_oct %>% 
  # Add placeholder rows for data gaps that are greater than 7 days to prevent
    # interpolation of large data gaps in the plot
  group_by(YearFct, Region) %>% 
  complete(Date = seq.Date(min(Date), max(Date), by = "day")) %>% 
  arrange(Date) %>% 
  mutate(
    na_val = is.na(NonDetect),
    na_val_run_total = sequence(rle(na_val)$lengths)
  ) %>% 
  filter(!(na_val == TRUE & na_val_run_total < 8)) %>% 
  ungroup() %>%
  select(!starts_with("na_val")) %>% 
  # Restructure data to long format
  select(-c(AvgCI, InvalidOrMissing)) %>% 
  pivot_longer(
    cols = NonDetect:VeryHigh,
    names_to = "CIcategory",
    values_to = "CIcount"
  ) %>% 
  # Apply factor order to the Cyano Index categories
  mutate(
    CIcategory = factor(
      CIcategory, 
      levels = ci_cat_levels, 
      labels = ci_cat_labels
    )
  ) %>% 
  # Convert NA values in the count column to zero to break interpolation of
    # large data gaps in the plot
  replace_na(list(CIcount = 0))
```

```{r create area plot, fig.height = 8}
# Create Area Plot
plt_hab_sat_area <-df_hab_sat_area_plt %>% 
  ggplot(aes(x = Date, y = CIcount, fill = CIcategory)) +
  geom_area(position = "fill") +
  facet_grid(
    rows = vars(Region),
    cols = vars(YearFct),
    scales = "free_x"
  ) +
  scale_x_date(
    name = "Date",
    breaks = breaks_pretty(5),
    labels = label_date_short(c(NA, "%b", "%d", NA)),
    expand = expansion(mult = 0.03)
  ) +
  scale_y_continuous(
    name = "Percent of valid pixels within each Cyano Index Category",
    labels = percent_format()
  ) +
  scale_fill_manual(
    name = "Cyano Index\nCategory",
    values = ci_cat_color_pal
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )

plt_hab_sat_area
```

## 2021 Maps

### Import and Prepare Satellite Data

From the area plots, there was a large bloom in Franks Tract during the summer of 2021. Let's take a closer look at the HAB satellite data during this time to see how the bloom appeared spatially. In order to do this, we will need to download the HAB satellite data from the SFEI website for July and August 2021 if that hasn't been done already. Then we'll need to prepare it to be used in the maps.

```{r download hab satellite data jul aug 2021}
# Download HAB satellite data for July and August 2021:
# Set download to TRUE if need to download harmful algal bloom (HAB) satellite data
download <- FALSE

# Download HAB satellite data to local computer if necessary
if (download == TRUE) {
  # Define subfolder directory to store .tif files
  dir_hab_sat <- here("EDB/Spatial_data")
  
  # Function to download and unzip harmful algal bloom (HAB) satellite data (cyanobacteria abundance)
    # from the https://fhab.sfei.org/ website
  download_hab <- function(hab_yr, hab_month) {
    hab_url <- glue("https://fhab.sfei.org/lib/download.php?request=download&dltype=month&year={hab_yr}&month={hab_month}&product=Mosaic")
    out_path <- file.path(dir_hab_sat, glue("mosaic_{hab_yr}_{hab_month}.zip"))
    
    curl_download(hab_url, out_path)
    unzip(out_path, exdir = dir_hab_sat)
    Sys.sleep(5)
  }
  
  # Download data for July and August 2021
  hab_2021 <- c(7, 8)
  for (i in hab_2021) {download_hab(2021, i)}
  
  # Remove .zip files
  invisible(file.remove(dir(path = dir_hab_sat, pattern = "zip$", full.names = TRUE)))
}
```

Let's import all the necessary spatial data. For the satellite imagery, we'll focus on the be beginning, peak, and end of the Cyano bloom in Franks Tract during the summer of 2021.

```{r import spatial data}
# Create a vector of all file paths for the spatial data data
fp_spt_data <- dir(here("EDB/Spatial_data"), full.names = TRUE)

# Import HAB satellite data for the beginning, peak, and end of the Cyano bloom
  # in Franks Tract in 2021:
# Pull out four .tif files representing the beginning, peak, and end of the Cyano bloom
fp_fr_bloom <- str_subset(fp_spt_data, "sentinel-3a.+(0710|0729|0810|0825)")

# Create a nested data frame to prepare the HAB satellite data for maps
df_hab_sat <-
  tibble(
    date_chr = c("July 10, 2021", "July 29, 2021", "Aug 10, 2021", "Aug 25, 2021"),
    fp = fp_fr_bloom,
    strs_prx_obj = map(fp, read_stars, proxy = TRUE)
  ) %>%
  select(-fp)

# Import the polygon shapefile for the four open water regions in the Delta
  # including Franks Tract
sf_ow_delta <- read_sf(str_subset(fp_spt_data, "Franks_Mildr_CCF_LibIsl.shp"))
```

Next, we'll need to transform the coordinate reference system (CRS) of the open water regions and WW_Delta shapefiles to the CRS of the HAB satellite data so they can all be plotted together on the maps. For the shapefile of open water regions, we'll only keep the polygon for Franks Tract since that's the region we are most interested in.

```{r prepare sf polygon data}
crs_hab_sat <- st_crs(df_hab_sat$strs_prx_obj[[1]])
sf_franks_32611 <- st_transform(sf_ow_delta, crs = crs_hab_sat) %>% filter(HNAME == "Franks Tract")
WW_Delta_32611 <- st_transform(WW_Delta, crs = crs_hab_sat)
```

We also need to create a bounding box of the Franks Tract shapefile which will be used to crop the satellite data and define the boundaries of the maps. We'll add a 2.5 km buffer to slightly expand the bounding box to ensure no desired data is removed.

```{r create bbox sf franks tract}
bbox_franks <- st_bbox(st_buffer(sf_franks_32611, 2500))
```

Lastly, we need to prepare the HAB satellite data for the maps.

```{r prepare hab satellite data}
df_hab_sat_c <- df_hab_sat %>%
  mutate(
    strs_obj_f =
      # Crop HAB satellite data to bounding box of the Franks Tract shapefile
      map(strs_prx_obj, ~st_crop(.x, bbox_franks) %>%
        # rename attribute to be more descriptive
        setNames("pixel_val") %>%
        # Convert pixel values to Cyano Index categories
        mutate(
          pixel_val = as.numeric(as.character(pixel_val)),
          pixel_val = case_when(
            pixel_val == 0 ~ "NonDetect",
            pixel_val <= 41 ~ "Low",
            pixel_val <= 99 ~ "Moderate",
            pixel_val <= 183 ~ "High",
            pixel_val <= 250 ~ "VeryHigh",
            TRUE ~ NA_character_
          ),
          pixel_val = factor(pixel_val, levels = ci_cat_levels, labels = ci_cat_labels)
        ) %>%
        # Convert to stars object
        st_as_stars()
    )
  ) %>%
  select(-strs_prx_obj)
```

### Create Maps

```{r hab satellite maps function}
# Function to create maps of HAB satellite data
create_hab_map <- function(strs_obj, map_title, x_txt_lab, y_txt_lab) {
  p <- ggplot() +
    geom_stars(data = strs_obj, na.rm = TRUE) +
    scale_fill_manual(
      name = "Cyano Index\nCategory",
      drop = FALSE,
      na.translate = FALSE,
      values = ci_cat_color_pal
    ) +
    geom_sf(
      data = WW_Delta_32611, 
      alpha = 0,
      color = "royalblue3",
      size = 0.3
    ) +
    geom_sf(
      data = sf_franks_32611, 
      alpha = 0, 
      color = "green3",
      size = 1.1
    ) +
    coord_sf(
      xlim = c(bbox_franks$xmin, bbox_franks$xmax),
      ylim = c(bbox_franks$ymin, bbox_franks$ymax)
    ) + 
    ggtitle(map_title) +
    ylab(NULL) +
    scale_x_continuous(
      name = NULL,
      breaks = seq(-121.68, -121.56, by = 0.04)
    ) +
    theme_bw()
  
  # Only include x-axis tick labels for the bottom most maps
  if (x_txt_lab == FALSE) {
    p <- p + theme(axis.text.x = element_blank())
  }
  
   # Only include y-axis tick labels for the left most maps
  if (y_txt_lab == FALSE) {
    p <- p + theme(axis.text.y = element_blank())
  }
  
  return(p)
}
```

```{r create hab satellite maps, fig.height = 7}
# Create maps of HAB satellite data for each day
map_hab_sat <- df_hab_sat_c %>% 
  mutate(
    x_txt = c(FALSE, FALSE, TRUE, TRUE),
    y_txt = c(TRUE, FALSE, TRUE , FALSE),
    hab_map = pmap(
      list(strs_obj_f, date_chr, x_txt, y_txt),
      create_hab_map
    )
  )

# Combine maps for each day into one
map_hab_sat_c <- wrap_plots(map_hab_sat$hab_map, ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")

map_hab_sat_c
```

## Export Figures

Export the area plot and maps of HAB satellite data for the June 2022 version of the HABs/Weeds report. These .jpg files were moved to the EDB SharePoint site.

```{r export figures, eval = FALSE}
# Export area plot
ggsave(
  here("EDB/CI_category_area_plot.jpg"),
  plot = plt_hab_sat_area,
  width = 6.5,
  height = 7.5,
  units = "in"
)

# Export map of HAB satellite data
ggsave(
  here("EDB/CI_category_map_2021.jpg"),
  plot = map_hab_sat_c,
  width = 6.5,
  height = 6.5,
  units = "in"
)
```

