---
title: "Survey Sampling Locations"
author: "Dave Bosworth"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = "")
```

# Purpose

Provide a map and a table summarizing the sampling effort of the various surveys that collect discrete nutrient and chlorophyll-a data. These will aid in the analyses for the Emergency Drought Barrier HABs/Weeds report.

The following interactive map and sampling effort table only includes stations where all three nutrients (dissolved ammonia, dissolved nitrate + nitrite, and dissolved ortho-phosphate) and chlorophyll-a have been collected during their periods of record from 2014-2021.


# Global code and functions

```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(lubridate)
# install.packages("devtools")
# devtools::install_github("mountaindboz/EDBdata", ref = "v0.3.0)
library(EDBdata)
library(sf)
library(leaflet)
library(here)
library(dataRetrieval)
library(htmltools)
library(knitr)
library(kableExtra)

# Check if we are in the correct working directory
i_am("EDB/nutr_chla_survey_summary.Rmd")
```


# Import and Prepare Data

```{r import data, message = FALSE}
# Pull out station coordinates from hab_nutr_chla_mvi data frame
df_coord_all <- hab_nutr_chla_mvi %>% distinct(Source, Station, Latitude, Longitude)

# Bring in Station Names for those that have that information readily available:
# USGS_CAWSC:
cawsc_sta <- df_coord_all %>% 
  filter(Source == "USGS_CAWSC") %>% 
  pull(Station)

df_cawsc_coord <- whatWQPsites(siteid = cawsc_sta)
  
# DWR_EMP:
df_emp_coord <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=edi.458.4&entityid=827aa171ecae79731cc50ae0e590e5af")
```

```{r clean and combine coordinates}
# Create data frame of Station Names to be joined to the station coordinates
df_sta_names <- df_cawsc_coord %>% 
  select(
    Station = MonitoringLocationIdentifier, 
    StationName = MonitoringLocationName
  ) %>% 
  mutate(Source = "USGS_CAWSC") %>% 
  bind_rows(df_emp_coord %>% select(Station, StationName = Location)) %>% 
  replace_na(list(Source = "DWR_EMP"))

# Join station names to station coordinates
df_coord_all_c <- df_coord_all %>% 
  left_join(df_sta_names, by = c("Source", "Station")) %>% 
  # Add DWR_NCRO Station Names
  mutate(
    StationName = case_when(
      Station == "FAL" ~ "False River near Oakley",
      Station == "BET" ~ "Bethel Island near Piper Slough",
      Station == "HOL" ~ "Holland Cut near Bethel Island",
      Station == "FCT" ~ "Fisherman's Cut",
      Station == "OSJ" ~ "Old River near Franks Tract",
      Station == "TSL" ~ "Three Mile Slough at San Joaquin River",
      TRUE ~ StationName
    )
  ) %>% 
  rename(StationCode = Station)
```

```{r convert coord df to sf}
# Convert the station coordinates data frame to an sf object (WGS84 datum)
sf_coord_all <- df_coord_all_c %>% st_as_sf(coords = c("Longitude", "Latitude"), crs = 4326)
```


# Sampling Locations Map

This interactive map allows for zooming and panning. Hover over a station marker to get more information about it.

```{r create leaflet map}
# Define color palette for Surveys 
color_pal <- colorFactor(palette = "viridis", domain = sf_coord_all$Source)

# Create a list for the CircleMarker labels to be used with HTML (allows for line breaks)
map_labels <- 
  pmap(
    list(
      sf_coord_all$Source,
      sf_coord_all$StationCode,
      sf_coord_all$StationName
    ),
    ~ paste0(
      "Source: ", ..1,
      "<br>Station Code: ", ..2,
      "<br>Station Name: ", ..3
    )
  )

# Create map using leaflet
leaflet(sf_coord_all) %>% 
  # Needed to use this base map since Standard OpenStreetMap wasn't showing correctly
  addProviderTiles("OpenStreetMap.DE") %>%
  addCircleMarkers(
    radius = 4,
    fillColor = ~color_pal(Source),
    fillOpacity = 1,
    weight = 0.5,
    color = "black",
    opacity = 1,
    label = map(map_labels, HTML)
  ) %>% 
  addLegend(
    position = "topright",
    pal = color_pal,
    values = ~Source,
    title = "Survey"
  )
```


# Sampling Effort

Summary of the sampling effort for each parameter across stations and years:

```{r count stations years}
hab_nutr_chla_mvi %>%
  select(!c(ends_with("_Sign"), Microcystis)) %>%
  pivot_longer(
    cols = c(starts_with("Diss"), Chlorophyll),
    names_to = "Parameter",
    values_to = "Value"
  ) %>%
  filter(!is.na(Value)) %>%
  mutate(Year = year(Date)) %>%
  count(Source, Station, Parameter, Year) %>%
  arrange(Year) %>%
  pivot_wider(names_from = Year, values_from = n) %>%
  complete(nesting(Source, Station), Parameter) %>%
  arrange(Source, Station, Parameter) %>% 
  kable(format = "html") %>%
  kable_styling(bootstrap_options = "striped", fixed_thead = TRUE)
```

