```{r}
# Title: Exploratory analysis of WQ data
# Purpose: determine best regressions to use for the drought synthesis project
# Author: Sarah Perry
# Contact: sarah.perry@water.ca.gov
```

```{r}
library(tidyverse)
library(lubridate)
library(emmeans)
library(DroughtData)
library(here)
source(here("Water_Quality/analysis_feb2022/nutr_analysis_funcs.R"))
source(here("Water_Quality/analysis_feb2022/nutr_graph_funcs.R"))
```

# ----
LONG TERM DATA ANALYSIS (DROUGHT)
# ----

```{r}
# raw data
df_wq <- raw_nutr_1975_2021
df_wq$DissAmmonia_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'
df_wq$DissNitrateNitrite_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'
df_wq$DissOrthophos_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'

# remove outlier vals
df_wq <- df_wq %>% filter(
  !c(DissAmmonia_Sign == '<' & DissAmmonia >= .2),
  !c(DissOrthophos_Sign == '<' & DissOrthophos > .15),
  !c(DissNitrateNitrite_Sign == '<' & DissNitrateNitrite >= 3)
  )

# set seed
seed <- 42

df_wq$DissAmmonia <- replace_rl(df_wq, 'DissAmmonia', check_rl_col = FALSE, seed = seed)
df_wq$DissNitrateNitrite <- replace_rl(df_wq, 'DissNitrateNitrite', check_rl_col = FALSE, seed = seed)
df_wq$DissOrthophos <- replace_rl(df_wq, 'DissOrthophos', check_rl_col = FALSE, seed = seed)
```

# Dissolved Ammonia
## Seasonal Averages for Entire Delta (DissAmmonia)
```{r}
df_am_seasonal <- drt_avg_data(df = df_wq, data_var = DissAmmonia, avg_type = 'season', month.na = 'relaxed')
df_am_seasonal <- drt_add_yr_assign(df_am_seasonal)
df_am_seasonal$DissAmmonia <- log10(df_am_seasonal$DissAmmonia)
df_am_seasonal$Season <- factor(df_am_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_am_seasonal <- df_am_seasonal %>% arrange(., YearAdj, Season)

df_am_seasonal$DissAmmonia_onelag <- lag(df_am_seasonal$DissAmmonia, n = 1L)

lm_seasonal <- lm(DissAmmonia ~ Drought + Season + DissAmmonia_onelag, data = df_am_seasonal)

# plts
plt_am_lt_seas_di <- emm_plotter(
  lm_seasonal,
  df_am_seasonal,
  analyte = 'DissAmmonia',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Ammonia)')

plt_am_lt_seas_seas <- emm_plotter(
  lm_seasonal,
  df_am_seasonal,
  analyte = 'DissAmmonia',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Ammonia)')
```

# Regional Averages for Entire Delta (DissAmmonia)
```{r}
df_am_regional <- drt_avg_data(df = df_wq, data_var = DissAmmonia, avg_type = 'region', month.na = 'relaxed')
df_am_regional <- drt_add_yr_assign(df_am_regional)
df_am_regional$DissAmmonia <- log10(df_am_regional$DissAmmonia)
df_am_regional$Region <- factor(df_am_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_am_regional <- df_am_regional %>% arrange(., Region, YearAdj)

df_am_regional <- df_am_regional %>%
  group_by(Region) %>%
  mutate(DissAM_onelag = dplyr::lag(DissAmmonia, n = 1, default = NA)) %>%
  ungroup()

lm_regional <- lm(DissAmmonia ~ Drought + Region + DissAM_onelag, data = df_am_regional)

# plts
plt_am_lt_reg_di <- emm_plotter(
  lm_regional,
  df_am_regional,
  analyte = 'DissAmmonia',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Ammonia)')

plt_am_lt_reg_reg <- emm_plotter(
  lm_regional,
  df_am_regional,
  analyte = 'DissAmmonia',
  grouping = 'Region',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Ammonia)')
```

# DissNitrateNitrite
## Seasonal Averages for Entire Delta
```{r message = FALSE}
df_nn_seasonal <- drt_avg_data(df = df_wq, data_var = DissNitrateNitrite, avg_type = 'season', month.na = 'relaxed')
df_nn_seasonal <- drt_add_yr_assign(df_nn_seasonal)
df_nn_seasonal$DissNitrateNitrite <- log10(df_nn_seasonal$DissNitrateNitrite)
df_nn_seasonal$Season <- factor(df_nn_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_nn_seasonal <- df_nn_seasonal %>% arrange(., YearAdj, Season)

df_nn_seasonal$DissNN_onelag <- lag(df_nn_seasonal$DissNitrateNitrite, n = 1L)

lm_seasonal <- lm(DissNitrateNitrite ~ Drought + Season + DissNN_onelag, data = df_nn_seasonal)

# plts
plt_nn_lt_seas_di <- emm_plotter(
  lm_seasonal,
  df_nn_seasonal,
  analyte = 'DissNitrateNitrite',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Nitrate/Nitrite)')

plt_nn_lt_seas_seas <- emm_plotter(
  lm_seasonal,
  df_nn_seasonal,
  analyte = 'DissNitrateNitrite',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Nitrate/Nitrite)')
```

## Regional Averages for the Entire Delta (DissNitrateNitrite)
```{r}
df_nn_regional <- drt_avg_data(df = df_wq, data_var = DissNitrateNitrite, avg_type = 'region', month.na = 'relaxed')
df_nn_regional <- drt_add_yr_assign(df_nn_regional)
df_nn_regional$DissNitrateNitrite <- log10(df_nn_regional$DissNitrateNitrite)
df_nn_regional$Region <- factor(df_nn_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_nn_regional <- df_nn_regional %>% arrange(., Region, YearAdj)

df_nn_regional <- df_nn_regional %>%
  group_by(Region) %>%
  mutate(DissNN_onelag = dplyr::lag(DissNitrateNitrite, n = 1, default = NA)) %>%
  ungroup()

lm_regional <- lm(DissNitrateNitrite ~ Drought + Region + DissNN_onelag, data = df_nn_regional)

# plts
plt_nn_lt_reg_di <- emm_plotter(
  lm_regional,
  df_nn_regional,
  analyte = 'DissNitrateNitrite',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Nitrate/Nitrite)')

plt_nn_lt_reg_reg <- emm_plotter(
  lm_regional,
  df_nn_regional,
  analyte = 'DissNitrateNitrite',
  grouping = 'Region',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Nitrate/Nitrite)')
```

# Dissolved Orthophosphate
## Seasonal Averages for Entire Delta (DissOrthophos)
```{r}
df_op_seasonal <- drt_avg_data(df = df_wq, data_var = DissOrthophos, avg_type = 'season', month.na = 'relaxed')
df_op_seasonal <- drt_add_yr_assign(df_op_seasonal)
df_op_seasonal$DissOrthophos <- log10(df_op_seasonal$DissOrthophos)
df_op_seasonal$Season <- factor(df_op_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_op_seasonal <- df_op_seasonal %>% arrange(., YearAdj, Season)

df_op_seasonal$DissOrthophos_onelag <- lag(df_op_seasonal$DissOrthophos, n = 1L)

lm_seasonal <- lm(DissOrthophos ~ Drought + Season + DissOrthophos_onelag, data = df_op_seasonal)

# plts
plt_op_lt_seas_di <- emm_plotter(
  lm_seasonal,
  df_op_seasonal,
  analyte = 'DissOrthophos',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Orthophosphate)')

plt_op_lt_seas_seas <- emm_plotter(
  lm_seasonal,
  df_op_seasonal,
  analyte = 'DissOrthophos',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Orthophosphate)')
```

# Regional Averages for the Entire Delta (DissOrthophos)
```{r}
df_op_regional <- drt_avg_data(df = df_wq, data_var = DissOrthophos, avg_type = 'region', month.na = 'relaxed')
df_op_regional <- drt_add_yr_assign(df_op_regional)
df_op_regional$DissOrthophos <- log10(df_op_regional$DissOrthophos)
df_op_regional$Region <- factor(df_op_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_op_regional <- df_op_regional %>% arrange(., Region, YearAdj)


df_op_regional <- df_op_regional %>%
  group_by(Region) %>%
  mutate(DissOP_onelag = dplyr::lag(DissOrthophos, n = 1, default = NA)) %>%
  ungroup()

lm_regional <- lm(DissOrthophos ~ Drought + Region  + DissOP_onelag, data = df_op_regional)

#plts
plt_op_lt_reg_di <- emm_plotter(
  lm_regional,
  df_op_regional,
  analyte = 'DissOrthophos',
  grouping = 'Drought',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Orthophosphate)')

plt_op_lt_reg_reg <- emm_plotter(
  lm_regional,
  df_op_regional,
  analyte = 'DissOrthophos',
  grouping = 'Region',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Orthophosphate)')

```

```{r}
# save plots
# dissam
bp_em_DissAmmonia_di_both <- dual_year_plts(plt_am_lt_reg_di, plt_am_lt_seas_di, 'Drought')
ggsave(
  filename = here("Water_Quality/bp_em_DissAmmonia_di_both.jpg"), 
  plot = bp_em_DissAmmonia_di_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)

# dissnn
bp_em_DissNitrateNitrite_di_both <- dual_year_plts(plt_nn_lt_reg_di, plt_nn_lt_seas_di, 'Drought')
ggsave(
  filename = here("Water_Quality/bp_em_DissNitrateNitrite_di_both.jpg"), 
  plot = bp_em_DissNitrateNitrite_di_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)

# dissorthophos
bp_em_DissOrthophos_di_both <- dual_year_plts(plt_op_lt_reg_di, plt_op_lt_seas_di, 'Drought')
ggsave(
  filename = here("Water_Quality/bp_em_DissOrthophos_di_both.jpg"), 
  plot = bp_em_DissOrthophos_di_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)
```

# ----
SHORT TERM DATA ANALYSIS (YEARS)
# ----

From Dave's Drought package
```{r}
# raw data
df_wq <- raw_nutr_2013_2021
df_wq$DissAmmonia_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'
df_wq$DissNitrateNitrite_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'
df_wq$DissOrthophos_Sign[df_wq$DissAmmonia_Sign == '< (estimated)'] <- '<'

# remove outlier vals
df_wq <- df_wq %>% filter(
  !c(DissAmmonia_Sign == '<' & DissAmmonia >= .2),
  !c(DissOrthophos_Sign == '<' & DissOrthophos > .15),
  !c(DissNitrateNitrite_Sign == '<' & DissNitrateNitrite >= 3)
  )

# set seed
seed <- 42

df_wq$DissAmmonia <- replace_rl(df_wq, 'DissAmmonia', seed = seed)
df_wq$DissNitrateNitrite <- replace_rl(df_wq, 'DissNitrateNitrite', seed = seed)
df_wq$DissOrthophos <- replace_rl(df_wq, 'DissOrthophos', seed = seed)
df_wq <- filter(df_wq, Region != 'Suisun Marsh')
```

# Dissolved Ammonia
## Seasonal Averages for Entire Delta (DissAmmonia)
```{r}
df_am_seasonal <- drt_avg_data(df = df_wq, data_var = DissAmmonia, avg_type = 'season', month.na = 'relaxed')
df_am_seasonal <- drt_add_yr_assign(df_am_seasonal)
df_am_seasonal$DissAmmonia <- log10(df_am_seasonal$DissAmmonia)
df_am_seasonal$YearAdj <- as.factor(df_am_seasonal$YearAdj)
df_am_seasonal$Season <- factor(df_am_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_am_seasonal <- df_am_seasonal %>% arrange(., YearAdj, Season)

lm_seasonal <- lm(DissAmmonia ~ YearAdj + Season, data = df_am_seasonal)

# plts
plt_am_st_seas_yr <- emm_plotter(
  lm_seasonal,
  df_am_seasonal,
  analyte = 'DissAmmonia',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.85,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Ammonia)')

plt_am_st_seas_seas <- emm_plotter(
  lm_seasonal,
  df_am_seasonal,
  analyte = 'DissAmmonia',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Ammonia)')

plt_am_st_seas_yr
```

# Regional Averages for Entire Delta (DissAmmonia)
```{r}
df_am_regional <- drt_avg_data(df = df_wq, data_var = DissAmmonia, avg_type = 'region', month.na = 'relaxed')
df_am_regional <- drt_add_yr_assign(df_am_regional)
df_am_regional$DissAmmonia <- log10(df_am_regional$DissAmmonia)
df_am_regional$YearAdj <- as.factor(df_am_regional$YearAdj)
df_am_regional$Region <- factor(df_am_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_am_regional <- df_am_regional %>% arrange(., Region, YearAdj)

lm_regional <- lm(DissAmmonia ~ YearAdj + Region, data = df_am_regional)

# plts
plt_am_st_reg_yr <- emm_plotter(
  lm_regional,
  df_am_regional,
  analyte = 'DissAmmonia',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.85,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Ammonia)')

plt_am_st_reg_reg <- emm_plotter(
  lm_regional,
  df_am_regional,
  analyte = 'DissAmmonia',
  grouping = 'Region',
  fatten = 0.8,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Ammonia)')
```

# DissNitrateNitrite
## Seasonal Averages for Entire Delta
```{r message = FALSE}
df_nn_seasonal <- drt_avg_data(df = df_wq, data_var = DissNitrateNitrite, avg_type = 'season', month.na = 'relaxed')
df_nn_seasonal <- drt_add_yr_assign(df_nn_seasonal)
df_nn_seasonal$DissNitrateNitrite <- log10(df_nn_seasonal$DissNitrateNitrite)
df_nn_seasonal$YearAdj <- as.factor(df_nn_seasonal$YearAdj)
df_nn_seasonal$Season <- factor(df_nn_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_nn_seasonal <- df_nn_seasonal %>% arrange(., YearAdj, Season)

lm_seasonal <- lm(DissNitrateNitrite ~ YearAdj + Season, data = df_nn_seasonal)

# plts
plt_nn_st_seas_yr <- emm_plotter(
  lm_seasonal,
  df_nn_seasonal,
  analyte = 'DissNitrateNitrite',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.85,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Nitrate/Nitrite)')

plt_nn_st_seas_seas <- emm_plotter(
  lm_seasonal,
  df_nn_seasonal,
  analyte = 'DissNitrateNitrite',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Nitrate/Nitrite)')
```

## Regional Averages for the Entire Delta (DissNitrateNitrite)
```{r}
df_nn_regional <- drt_avg_data(df = df_wq, data_var = DissNitrateNitrite, avg_type = 'region', month.na = 'relaxed')
df_nn_regional <- drt_add_yr_assign(df_nn_regional)
df_nn_regional$DissNitrateNitrite <- log10(df_nn_regional$DissNitrateNitrite)
df_nn_regional$YearAdj <- as.factor(df_nn_regional$YearAdj)
df_nn_regional$Region <- factor(df_nn_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_nn_regional <- df_nn_regional %>% arrange(., Region, YearAdj)

lm_regional <- lm(DissNitrateNitrite ~ YearAdj + Region, data = df_nn_regional)

# plts
plt_nn_st_reg_yr <- emm_plotter(
  lm_regional,
  df_nn_regional,
  analyte = 'DissNitrateNitrite',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.85,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Nitrate/Nitrite)')

plt_nn_st_reg_reg <- emm_plotter(
  lm_regional,
  df_nn_regional,
  analyte = 'DissNitrateNitrite',
  grouping = 'Region',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Nitrate/Nitrite)')
```

# Dissolved Orthophosphate
## Seasonal Averages for Entire Delta (DissOrthophos)
```{r}
df_op_seasonal <- drt_avg_data(df = df_wq, data_var = DissOrthophos, avg_type = 'season', month.na = 'relaxed')
df_op_seasonal <- drt_add_yr_assign(df_op_seasonal)
df_op_seasonal$DissOrthophos <- log10(df_op_seasonal$DissOrthophos)
df_op_seasonal$YearAdj <- as.factor(df_op_seasonal$YearAdj)
df_op_seasonal$Season <- factor(df_op_seasonal$Season, levels = c('Winter', 'Spring', 'Summer', 'Fall'))
df_op_seasonal <- df_op_seasonal %>% arrange(., YearAdj, Season)

lm_seasonal <- lm(DissOrthophos ~ YearAdj + Season, data = df_op_seasonal)

# plts
plt_op_st_seas_yr <- emm_plotter(
  lm_seasonal,
  df_op_seasonal,
  analyte = 'DissOrthophos',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.85,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Orthophosphate)')

plt_op_st_seas_seas <- emm_plotter(
  lm_seasonal,
  df_op_seasonal,
  analyte = 'DissOrthophos',
  grouping = 'Season',
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Orthophosphate)')
```

# Regional Averages for the Entire Delta (DissOrthophos)
```{r}
df_op_regional <- drt_avg_data(df = df_wq, data_var = DissOrthophos, avg_type = 'region', month.na = 'relaxed')
df_op_regional <- drt_add_yr_assign(df_op_regional)
df_op_regional$DissOrthophos <- log10(df_op_regional$DissOrthophos)
df_op_regional$YearAdj <- as.factor(df_op_regional$YearAdj)
df_op_regional$Region <- factor(df_op_regional$Region, levels = c('Suisun Bay', 'Confluence', 'SouthCentral', 'North'))
df_op_regional <- df_op_regional %>% arrange(., Region, YearAdj)

lm_regional <- lm(DissOrthophos ~ YearAdj + Region, data = df_op_regional)

#plts
plt_op_st_reg_yr <- emm_plotter(
  lm_regional,
  df_op_regional,
  analyte = 'DissOrthophos',
  grouping = 'YearAdj',
  fill = 'Drought',
  fill_type = 'rect',
  rect_gap = 0.02,
  fatten = 0.8,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  drt_color_pal_drought() +
  ylab('log(Dissolved Orthophosphate)')

plt_op_st_reg_reg <- emm_plotter(
  lm_regional,
  df_op_regional,
  analyte = 'DissOrthophos',
  grouping = 'Region',
  fatten = 0.8,
  line_size = .7,
  text_size = 6
) +
  theme_bw() +
  ylab('log(Dissolved Orthophosphate)')
```

```{r}
# save plots
# dissam
bp_em_DissAmmonia_yt_both <- dual_year_plts(plt_am_st_reg_yr, plt_am_st_seas_yr, 'Year', angle = 45)
ggsave(
  filename = here("Water_Quality/bp_em_DissAmmonia_yt_both.jpg"), 
  plot = bp_em_DissAmmonia_yt_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)

# dissnn
bp_em_DissNitrateNitrite_yt_both <- dual_year_plts(plt_nn_st_reg_yr, plt_nn_st_seas_yr, 'Year', angle = 45)
ggsave(
  filename = here("Water_Quality/bp_em_DissNitrateNitrite_yt_both.jpg"), 
  plot = bp_em_DissNitrateNitrite_yt_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)

# dissorthophos
bp_em_DissOrthophos_yt_both <- dual_year_plts(plt_op_st_reg_yr, plt_op_st_seas_yr, 'Year', angle = 45)
ggsave(
  filename = here("Water_Quality/bp_em_DissOrthophos_yt_both.jpg"), 
  plot = bp_em_DissOrthophos_yt_both, 
  width = 8, 
  height = 3.5,
  units = "in"
)
```

# ----
SAVE AS FOUR PLOTS
# ----
```{r}
plt_am_st_seas_yr
```

```{r}
# seasonal
#dissam
bp_am_seasonal <- quad_year_plts(plt_am_lt_seas_di, plt_am_st_seas_yr, plt_am_lt_seas_seas, plt_am_st_seas_seas,
                                 type = 'Season', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_am_seasonal.jpg"),
  plot = bp_am_seasonal, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)

#dissam
bp_nn_seasonal <- quad_year_plts(plt_nn_lt_seas_di, plt_nn_st_seas_yr, plt_nn_lt_seas_seas, plt_nn_st_seas_seas,
                                 type = 'Season', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_nn_seasonal.jpg"),
  plot = bp_nn_seasonal, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)

#dissam
bp_op_seasonal <- quad_year_plts(plt_op_lt_seas_di, plt_op_st_seas_yr, plt_op_lt_seas_seas, plt_op_st_seas_seas,
                                 type = 'Season', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_op_seasonal.jpg"),
  plot = bp_op_seasonal, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)
```

```{r}
# regional
#dissam
bp_am_regional <- quad_year_plts(plt_am_lt_reg_di, plt_am_st_reg_yr, plt_am_lt_reg_reg, plt_am_st_reg_reg,
                                 type = 'Region', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_am_regional.jpg"),
  plot = bp_am_regional, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)

#dissam
bp_nn_regional <- quad_year_plts(plt_nn_lt_reg_di, plt_nn_st_reg_yr, plt_nn_lt_reg_reg, plt_nn_st_reg_reg,
                                 type = 'Region', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_nn_regional.jpg"),
  plot = bp_nn_regional, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)

#dissam
bp_op_regional <- quad_year_plts(plt_op_lt_reg_di, plt_op_st_reg_yr, plt_op_lt_reg_reg, plt_op_st_reg_reg,
                                 type = 'Region', txt_size = 15, label_size = 17)
ggsave(
  here("Water_Quality/bp_op_regional.jpg"),
  plot = bp_op_regional, 
  width = 11.5, 
  height = 9.5,
  units = "in"
)
```

```{r}
plt_am_st_reg_yr
```

All exported figures were moved to the Drought Synthesis SharePoint site

